---
title: "Untitled"
output: html_document
---

# 1. Import arrow package

The arrow package is already installed, so no need to install it. Just import it.

```{r}
# import the package
library(arrow)
```

# 2. Read raw data to a dataframe
1. Create minio virtual filesystem

```{r}
minio <- S3FileSystem$create(
   access_key = Sys.getenv("AWS_ACCESS_KEY_ID"),
   secret_key = Sys.getenv("AWS_SECRET_ACCESS_KEY"),
   session_token = Sys.getenv("AWS_SESSION_TOKEN"),
   scheme = "https",
   endpoint_override = Sys.getenv("AWS_S3_ENDPOINT")
   )
```

2. Read raw data

```{r}
raw_data_path <- "pengfei/diffusion/data_format/sf_fire/parquet/arrow_sf_fire_gzip/546fa4b65a6a4f1f983f31f0eaf586c4.parquet"

df_sf_fire <- read_parquet(minio$path(raw_data_path))
head(df_sf_fire)
```

# 3. Write parquet to s3 benchmark

We will use the write_parquet function to benchmark the write time

write_parquet(
  x,
  sink,
  chunk_size = NULL,
  version = NULL,
  compression = default_parquet_compression(),
  compression_level = NULL,
  use_dictionary = NULL,
  write_statistics = NULL,
  data_page_size = NULL,
  use_deprecated_int96_timestamps = FALSE,
  coerce_timestamps = NULL,
  allow_truncated_timestamps = FALSE,
  properties = NULL,
  arrow_properties = NULL
)

```{r}
devtools::install_github("collectivemedia/tictoc")

library(tictoc)
```

## 3.1 Write parquet without compression
```{r}
# write parquet as uncompressed
start_time <- Sys.time()
output_path="pengfei/diffusion/data_format/sf_fire/parquet/r_arrow/r_arrow_sf_fire_none"
write_parquet(df_sf_fire, sink = minio$path(output_path), compression = "uncompressed")
end_time <- Sys.time()
latency <- end_time-start_time
cat("Latency for write parquet without compression: ", latency)

```

## 3.2 Write parquet with snappy

```{r}
# write parquet with snappy compression
start_time <- Sys.time()
output_path="pengfei/diffusion/data_format/sf_fire/parquet/r_arrow/r_arrow_sf_fire_snappy.parquet"
write_parquet(df_sf_fire, sink = minio$path(output_path), compression = "snappy")
end_time <- Sys.time()
latency <- end_time-start_time
cat("Latency for write parquet with snappy compression: ", latency)
```

## 3.3 Write parquet with gzip

```{r}

# write parquet as gzip
# start_time <- Sys.time()
tic()
output_path="pengfei/diffusion/data_format/sf_fire/parquet/r_arrow/r_arrow_sf_fire_gzip.parquet"
write_parquet(df_sf_fire, sink = minio$path(output_path), compression = "gzip")
toc()
# end_time <- Sys.time()
# latency <- end_time-start_time
# cat("Latency for write parquet with gzip compression: ", latency)
```


## 3.3 Write parquet with lz4

```{r}
# write parquet as lz4
tic()
output_path="pengfei/diffusion/data_format/sf_fire/parquet/r_arrow/r_arrow_sf_fire_lz4.parquet"
write_parquet(df_sf_fire, sink = minio$path(output_path), compression = "lz4")
toc()
```

## 3.4 Write parquet with zstd

```{r}
# write parquet as zstd
tic()
output_path="pengfei/diffusion/data_format/sf_fire/parquet/r_arrow/r_arrow_sf_fire_zstd.parquet"
write_parquet(df_sf_fire, sink = minio$path(output_path), compression = "zstd")
toc()
```
## 3.5 Write parquet with brotli

```{r}
# write parquet as brotli
tic()
output_path="pengfei/diffusion/data_format/sf_fire/parquet/r_arrow/r_arrow_sf_fire_brotli.parquet"
write_parquet(df_sf_fire, sink = minio$path(output_path), compression = "brotli")
toc()
```


