---
title: "Read parquet files from s3 by using Arrow"
output: html_document
---

# Read parque file from s3 by using Arrow 

## Step 1. Install the arrow package with s3 support

By default, for linux the S3 support is not enabled in the default build. To enable it, set the environment variable **LIBARROW_MINIMAL=false** or **NOT_CRAN=true** to choose the full-featured build

You can find the original doc [here](https://arrow.apache.org/docs/r/articles/install.html)
```{r}
# don’t use mini when compile arrow
Sys.setenv(LIBARROW_MINIMAL="false")

# Download and compile the package, it may take a while. Please be patient
install.packages("arrow", repos = "https://cloud.r-project.org")

# import the package
library(arrow)

```

## Step 2. Configure a s3 file system

We know this is not the only way, you can also use directly the s3 uri **s3://[access_key:secret_key:session_token@]bucket/path[?region=]**.

But we find the file system solution is easier to use. Below code will create
a virtual filesystem called minio.  

```{r}
minio <- S3FileSystem$create(
   access_key = Sys.getenv("AWS_ACCESS_KEY_ID"),
   secret_key = Sys.getenv("AWS_SECRET_ACCESS_KEY"),
   session_token = Sys.getenv("AWS_SESSION_TOKEN"),
   scheme = "https",
   endpoint_override = Sys.getenv("AWS_S3_ENDPOINT")
   )
```

You can have **multiple virtual filesystem** at the same time. For example, below create
a file system that links to DigitalOcean s3 storage. You can find 
NYC Taxi dataset in this location.

```{r}

taxi <- S3FileSystem$create(
  anonymous = TRUE,
  scheme = "https",
  endpoint_override = "sfo3.digitaloceanspaces.com"
)


```

## Step 3. Explore the data

Arrow provides you two options to use the data from the s3
1. Download it from remote server to local file system
2. Read the data into memory and return a dataframe

### 3.1 Download the data to local file system

We can use function **copy_files** to download the files. Note, you need to 
change the path of the first example. 


```{r}

# copy the file from s3 to local file
# change the below path to a data path which is in your bucket 
data_path <- "pengfei/diffusion/data_format/sf_fire/parquet/raw"
# change the below path, if you want to store the data somewhere else
local_data_path <- "~/sf_fire/raw/"

try(dir.create(local_data_path, recursive = T))
arrow::copy_files(minio$path(data_path), local_data_path )

```


```{r}
# copy the data of January 2009 to local file
try(dir.create("~/nyc-taxi/2009/01", recursive = T))

arrow::copy_files(taxi$path("nyc-taxi/2009/01"), "~/nyc-taxi/2009/01")

```


After the download, you can check the file by using the Terminal.

```{bash}
ls ~/sf_fire/raw

ls ~/nyc-taxi/2009/01
```

### 3.2 Read the file from local file system

As the files are downloaded to the local file system, we can read it easily by using
the function **read_parquet**.

```{r}

df_local <- read_parquet("~/sf_fire/raw/part-00000-0f7e798f-924f-4d08-8196-73fd17a47c4e-c000.gz.parquet")

head(df_local)

```

### 3.3 Read data into memory and return a datafame 

We can also directly load the remote data into memory and use it as a data frame

```{r}
# change the below path to a parquet file that is in your bucket
remote_parquet_file <- "pengfei/diffusion/data_format/sf_fire/parquet/raw/part-00000-0f7e798f-924f-4d08-8196-73fd17a47c4e-c000.gz.parquet"

df_remote <- read_parquet(minio$path(remote_parquet_file))

head(df_remote)

```

### 3.4 Read partitioned parquet file as dataset

If your parquet files are partitioned into small parquet files, you can read it by 
using the function **open_dataset**

```{r}

# use a local path and give a partition
nyc_path <- "~/nyc-taxi"
df_nyc <- open_dataset(
  nyc_path,
  partitioning = c("year", "month")
)
head(df_nyc)
```
You can also read the data from a remote path

```{r}

df_sf_fire <- open_dataset(
  minio$path(data_path)
)
head(df_sf_fire)
```



